The Second Data Release of the INT Photometric H$\alpha$ Survey of the Northern Galactic Plane (IPHAS DR2)

### 2014-04-16 paper/iphas-dr2.tex
Polishing Sec 4

(4) \emph{Photometric repeatability.} 
The IPHAS field-pair observing strategy normally
ensures that every pointing is immediately followed 
by an offset pointing at a displacement
of $+$5~arcmin in Dec and $+$5~arcmin in RA.
This allows pairs of images to be checked 
for the presence of clouds or electronic noise.
To exploit this information,
the overlap regions of all field pairs were systematically cross-matched
to verify the consistency of the photometry
for stars observed in both pointings.
We automatically rejected field pairs
in which more than 2 per cent of the stars 
showed an inconsistent measurement at the level of 0.2~mag,
or more than 26 per cent were inconsistent at the level of 0.1~mag.
These limits were set empirically after inspecting
the images and photometry by eye.

@@

For the purpose of providing an initial calibration 
of the $r$ and $i$ broadband fluxes,
photometric standard fields were observed every night.
The standards were chosen from a list based on 
the \cite{Landolt1992} and Stetson (http://cadcwww.dao.nrc.ca/standards) 
objects.
Two or three standard fields were observed 
during the evening and morning twilight,
and at intervals of 2-3 hours throughout the night.
The CASU pipeline automatically identified the observed standards 
and used them to determine a sigma-clipped average zeropoint \textsc{magzpt}
for each night and filter,
such that the number counts $DN$ 
in the pipeline-corrected CCD frames
relate to a magnitude $m$ as:
\begin{equation}
\begin{split}
   m  = & \textsc{magzpt} - 2.5 \log_{10}( DN / \textsc{exptime} ) \\
 &  - \textsc{extinct}\cdot(\textsc{airmass}-1) - \textsc{apcor} - \textsc{percorr},
\label{eqn:mag}
\end{split}
\end{equation}
where \textsc{exptime} is the exposure time in seconds,
\textsc{extinct} is the atmospheric extinction coefficient 
(set in the pipeline at 0.09 for $r$ and 0.05 for $i$ as representative
averages for La Palma),
\textsc{airmass} is the normalised optical path length 
through the atmosphere and
\textsc{apcor} is a correction for the flux
lost outside of the aperture used
(we adopt a 2\farcs 3-diameter circular aperture by default).
Finally, \textsc{percorr} is a correction based on the difference
between the median dark sky for a CCD against the median for all the CCDs, 
and as such is an ancillary correction 
to account for sporadic gain variations;
it is zero for most exposures.
All these quantities correspond to header keywords in the 
multi-extension FITS files produced by the CASU pipeline.

The broadband zeropoints were determined such that the resulting magnitude system
refers to the spectral energy distribution (SED) of Vega 
as the zero colour object. 
Colour equations were used to transform between the IPHAS passbands 
and the Johnson-Cousins system 
of the published standard star photometry.
The entire procedure has been found to deliver zeropoints which 
are accurate at the level of 1--2 per cent 
in stable photometric conditions \citep{Gonzalez-Solares2011}.

@@

\begin{table}
    \caption{Mean monochromatic flux of Vega in the IPHAS filter system,
             defined as 
             $\langle f_\lambda\rangle~=~\int{f_\lambda(\lambda) S(\lambda) \lambda d \lambda} / \int{S(\lambda)\lambda d \lambda}$,
             where $S(\lambda)$ is the photon response function 
             (which includes atmospheric transmission,
             filter transmission
             and CCD response)
             and $f_\lambda(\lambda)$ is the model SED for Vega
             by Kurucz (http://kurucz.harvard.edu/stars.html).
             For reference, we also provide the 
             filter equivalent width EW~=~$\int S(\lambda) d\lambda$,
             the mean photon wavelength 
             $\lambda_0~=~\int{S(\lambda)\lambda d\lambda} / \int{S(\lambda)d\lambda}$,
             and the pivot wavelength
             $\lambda_p~=~\sqrt{\int{S(\lambda)\lambda d\lambda} / \int{\frac{S(\lambda)}{\lambda} d\lambda}}$.
             These notations follow the definitions by \citet{Bessell2012}.
             After multiplying $\langle f_\lambda\rangle$ by the EW,
             we find that the detected flux for Vega in \ha\ is 3.14~magnitudes less
             than that received in $r$, 
          }
    \label{tbl:flux}
    \begin{center}
    \begin{tabular}{lccrc}
    \toprule
     Filter  & $\langle f_\lambda \rangle$  & \multicolumn{1}{c}{EW} &$\lambda_0$ & $\lambda_p$    \\
             & [erg\,cm$^{-2}$\,s$^{-1}$\,\AA$^{-1}$] & [\AA] & [\AA] & [\AA] \\
    \midrule
$r$ &  $2.47 \times 10^{-9}$  &  785.6 &  6223  &   6211   \\
\ha &  $1.81 \times 10^{-9}$ &   59.6 &  6568  &   6568    \\
$i$  &  $1.30 \times 10^{-9}$  &  759.9  &  7674  &   7674   \\
    \bottomrule
    \end{tabular}
    \end{center}
\end{table}

@@

Despite the best efforts made to obtain a nightly calibration,
large surveys naturally possess field-to-field variations
due to atmospheric changes during the night
and imperfections in the pipeline or the instrument
(e.g. the WFC is known to suffer from sporadic errors
in the timing of exposures).
This is demonstrated in Fig.~\ref{fig:ccd_before},
where we show the combined colour-colour diagram
for the set of 2\,801 quality-approved fields
which are located towards the Galactic anti-centre
($160\degr < \ell < 200\degr$).
Despite the fact that all the fields included in the diagram
have been quality-approved,
the main locus of stars is poorly defined
due to the presence of calibration errors,
which regularly exceed 0.1~mag.
For comparison, Fig.~\ref{fig:ccd_after}
shows the identical diagram
obtained after re-calibrating.

>>>>

(4) \emph{Photometric repeatability.} 
The IPHAS field-pair observing strategy normally
ensures that every pointing is immediately followed 
by an offset pointing at a displacement
of $+$5~arcmin in Dec and $+$5~arcmin in RA.
This allows pairs of images to be checked 
for the presence of clouds or electronic noise.
To exploit this information,
the overlap regions of all field pairs were systematically cross-matched
to verify the consistency of the photometry
for stars observed in both pointings.
We automatically rejected field pairs
in which more than 2 per cent of the stars 
showed an inconsistent measurement at the level of 0.2~mag,
or more than 25 per cent were inconsistent at the level of 0.1~mag.
These limits were set empirically after inspecting
the images and photometry by eye.

@@

For the purpose of providing an initial calibration 
of the $r$ and $i$ broadband fluxes,
photometric standard fields were observed every night.
The standards were chosen from a list based on 
the \cite{Landolt1992} and Stetson (http://cadcwww.dao.nrc.ca/standards) 
objects.
Two or three standard fields were observed 
during the evening and morning twilight,
and at intervals of 2-3 hours throughout the night.
The CASU pipeline automatically identified the observed standards 
and used them to determine a sigma-clipped average zeropoint \textsc{magzpt}
for each night and filter,
such that the number counts $DN$ 
in the pipeline-corrected CCD frames
relate to a magnitude $m$ as:
\begin{equation}
\begin{split}
   m  = & \textsc{magzpt} - 2.5 \log_{10}( DN / \textsc{exptime} ) \\
 &  - \textsc{extinct}\cdot(\textsc{airmass}-1) - \textsc{apcor} - \textsc{percorr},
\label{eqn:mag}
\end{split}
\end{equation}
where \textsc{exptime} is the exposure time in seconds,
\textsc{extinct} is the atmospheric extinction coefficient 
(set in the pipeline at 0.09 for $r$ and 0.05 for $i$
as representative averages for La Palma),
\textsc{airmass} is the normalised optical path length 
through the atmosphere,
and \textsc{apcor} is a correction for the flux
lost outside of the aperture used.
Finally, \textsc{percorr} is a correction based on the difference
between the median dark sky for a CCD against the median for all the CCDs, 
and as such is an ancillary correction 
to account for sporadic gain variations;
it is zero for most exposures.
All these quantities correspond to header keywords in the 
multi-extension FITS files produced by the CASU pipeline.

The broadband zeropoints were determined such that the resulting magnitude system
refers to the Spectral Energy Distribution (SED) of Vega 
as the zero colour object. 
Colour equations were used to transform between the IPHAS passbands 
and the Johnson-Cousins system 
of the published standard star photometry.
The entire procedure has been found to deliver zeropoints which 
are accurate at the level of 1--2 per cent 
in stable photometric conditions \citep{Gonzalez-Solares2011}.

@@

\begin{table}
    \caption{Mean monochromatic flux of Vega in the IPHAS filter system,
             defined as 
             $\langle f_\lambda\rangle~=~\int{f_\lambda(\lambda) S(\lambda) \lambda d \lambda} / \int{S(\lambda)\lambda d \lambda}$,
             where $S(\lambda)$ is the photon response function 
             (which includes atmospheric transmission,
             filter transmission
             and CCD response)
             and $f_\lambda(\lambda)$ is the model SED for Vega
             by Kurucz (http://kurucz.harvard.edu/stars.html).
             For reference, we also provide the 
             filter equivalent width EW~=~$\int S(\lambda) d\lambda$,
             the mean photon wavelength 
             $\lambda_0~=~\int{S(\lambda)\lambda d\lambda} / \int{S(\lambda)d\lambda}$,
             and the pivot wavelength
             $\lambda_p~=~\sqrt{\int{S(\lambda)\lambda d\lambda} / \int{\frac{S(\lambda)}{\lambda} d\lambda}}$.
             These notations follow the definitions by \citet{Bessell2012}.
             When multiplying $\langle f_\lambda\rangle$ by EW,
             we find that the flux received for Vega
             is 3.14~mag less in \ha\ than in $r$. 
          }
    \label{tbl:flux}
    \begin{center}
    \begin{tabular}{lccrc}
    \toprule
     Filter  & $\langle f_\lambda \rangle$  & \multicolumn{1}{c}{EW} &$\lambda_0$ & $\lambda_p$    \\
             & [erg\,cm$^{-2}$\,s$^{-1}$\,\AA$^{-1}$] & [\AA] & [\AA] & [\AA] \\
    \midrule
$r$ &  $2.47 \times 10^{-9}$  &  785.6 &  6223  &   6211   \\
\ha &  $1.81 \times 10^{-9}$ &   59.6 &  6568  &   6568    \\
$i$  &  $1.30 \times 10^{-9}$  &  759.9  &  7674  &   7674   \\
    \bottomrule
    \end{tabular}
    \end{center}
\end{table}

@@

Despite the best efforts made to obtain a nightly calibration,
large surveys naturally possess field-to-field variations
due to atmospheric changes during the night
and imperfections in the pipeline or the instrument
(e.g. the WFC is known to suffer from sporadic errors
in the timing of exposures).
This is demonstrated in Fig.~\ref{fig:ccd_before},
where we show the combined colour-colour diagram
for all of the 2,801 quality-approved fields
which are located towards the Galactic anti-centre
($160\degr < \ell < 200\degr$).
The main locus of stars is poorly defined
in the diagram due to the presence of 
fields which suffer from calibration errors,
which can exceed 0.1~mag or more.
For comparison, Fig.~\ref{fig:ccd_after}
shows the identical diagram
constructed after re-calibrating.
